Model

### 2019-08-22 stats/kids_combination.Rmd
modified model predictions

ggplot(data = p2, aes(x = age_num, y = correct, col = familiarObject, fill = familiarObject)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  geom_smooth(method = "loess", se = F, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(condition~familiarObject)+
  facet_grid(~condition)+
  #ggtitle("Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")+
  scale_fill_viridis_d(name = "familiar object")
```

>>>>

ggplot(data = p2, aes(x = age_num, y = correct, col = familiarObject, fill = familiarObject)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  #geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F, alpha = .5)+
  geom_smooth(method = "loess", se = F, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(condition~familiarObject)+
  facet_grid(~condition)+
  #ggtitle("Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")+
  scale_fill_viridis_d(name = "familiar object")
```

### 2019-08-22 stats/model.Rmd
modified model predictions

var lexicon1 = function(utterance, obj, sem_knowledge){
utterance.label == "novel_word" ? obj.shape == "novel_object" :
utterance.label == "familiar_word" ? flip(sem_knowledge) ? obj.shape == "familiar_object" :  obj.shape == "novel_object" : 
true
}

var lexicon2 = function(utterance, obj, sem_knowledge){
utterance.label == "novel_word" ? obj.shape == "familiar_object" :
utterance.label == "familiar_word" ? flip(sem_knowledge) ? obj.shape == "familiar_object" :  obj.shape == "novel_object" : 
true
}

@@

```{r rsa model}
rsaModel <- '
var literalListener = cache(function(utterance, priorProbs, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior); 
    var lexicon = lexiconObject[lexiconName];
    var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
    if ("label" in utterance) {
      var truthValue = lexicon(utterance, obj, sem_knowledge);
      condition(truthValue)
    }
    return obj.shape 
  }})
}, 10000)

@@

  var speakerOptimalityParameters = {
    intercept: uniformDrift({a: -2, b: 2, width: 0.5}),
    slope: uniformDrift({a: 0, b: 1, width: 0.5})
  }

@@

```{r}
me_model_data <- read_csv("../data/child_basic_me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(age = age_num - min(age_num))

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 10,
  cores = 10,
  inference_opts = list(method = "incrementalMH",
                        samples = 4000,
                        burn = 2000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 1000)
)

#saveRDS(sem_know_model, "../saves/sem_know_model.rds")

# 3 chains, 6000 iterations per chain, 2000 burn in, lag = 1
sem_know_model <- readRDS("../saves/run_correlations.rds")%>%
  filter(Chain != "15",
         Chain != "1",
         Chain != "30",
         Chain != "6",
         Chain != "5",
         Chain != "7",
         Chain != "3")

@@

@@

```{r}
me_model_data <- read_csv("../data/child_basic_me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(age = age_num - min(age_num))

sem_know_run_summary <- item_params%>%
  spread(parameter,value)%>%
  group_by(Chain,item)%>%
  summarise(overall_int = estimate_mode(int),
            overall_slope = estimate_mode(slope))

sem_know_run_iteration <- item_params%>%
  spread(parameter,value)

sem_know_chain <- sem_know_run_iteration %>%
  left_join(sem_know_run_summary)

chain_comp <- sem_know_chain %>%
  left_join(expand.grid(
      age = unique(me_model_data$age),
      item = unique(me_model_data$familiar)), .) %>%
  mutate(sem_know_ci = plogis(int + slope * age),
         sem_know = plogis(overall_int + overall_slope * age)) %>%
  group_by(Chain,item, age) %>%
  summarise(sem_know = estimate_mode(sem_know), 
            sem_know_uci = hdi_upper(sem_know_ci),
            sem_know_lci = hdi_lower(sem_know_ci))

ggplot(chain_comp, aes(x = age+2, y= sem_know, fill = factor(Chain)))+
  geom_line(size = .3, aes(col = factor(Chain)))+
  geom_ribbon(aes(ymin=sem_know_lci, ymax = sem_know_uci), alpha = .05)+
  ylab("Semantic knowledge")+
  xlab("Age")+
  ylim(0,1)+
  facet_wrap(~item, nrow = 3)+
  theme_few()+
  scale_colour_discrete()+
  scale_fill_discrete()

ggsave("../graphs/model_by_chain.pdf", width = 5, height = 4, scale = 1.5)
```

@@

so_run_iteration <- global_params%>%
  spread(parameter, value)

so_run_summary <- so_run_iteration%>%
  group_by(Chain)%>%
  summarise(overall_int = estimate_mode(int),
            overall_slope = estimate_mode(slope))

so_run <- so_run_iteration %>%
  left_join(so_run_summary)

so_run_comp <-so_run%>%
  group_by(Chain,Iteration)%>%
  expand(int,slope,overall_int,overall_slope,age = unique(me_model_data$age))%>%
  mutate(so_ci = int + slope * age,
         so = overall_int + overall_slope * age) %>%
  group_by(Chain, age) %>%
  summarise(so = estimate_mode(so), 
            so_uci = hdi_upper(so_ci),
            so_lci = hdi_lower(so_ci))

ggplot(so_run_comp, aes(x = age+2, y= so, fill = factor(Chain)))+
  geom_line(size = .4, aes(col = factor(Chain)))+
  geom_ribbon(aes(ymin=so_lci, ymax = so_uci), alpha = .05)+
  ylab("speaker optimality")+
  xlab("Age")+
  #facet_grid(~run)+
  theme_few()+
  scale_colour_discrete()+
  scale_fill_discrete()

#ggsave("../graphs/so_by_chain.pdf", width = 4, height = 3, scale = 1.5)
```

@@

sem_know_cor_params  <- readRDS("../saves/run_correlations.rds")%>%
  separate(Parameter, into = c("parameter", "type","item"), sep = ",")%>%
  filter(parameter == "parameter", type == "items")%>%
  separate(value, into = c("int", "slope"),sep = ", ")%>%
  mutate(int = as.numeric(str_remove_all(int, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]")))%>%
  gather(parameter, value, int, slope)

sem_know_cor_params <- sem_know_cor_params%>%
  filter(Chain != "15",
         Chain != "1",
         Chain != "30",
         Chain != "6",
         Chain != "5",
         Chain != "7",
         Chain != "3")

@@

```{r}
ggplot(sem_know_cor_params%>%filter(parameter == "int"), aes(x = value))+
  geom_density(alpha = 0.3)+
  xlab("Parameter")+
  facet_grid(Chain~item, scales = 'free')+
  theme_few()

@@

```{r}
item_params_cor <- sem_know_cor_params %>%
  group_by(Chain,item, parameter)%>%
  summarise(mode = estimate_mode(value))
```

@@

## For comparison: GLMM for ME effect and data
```{r}
me_data <- read_csv("../data/child_basic_me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(z_age = age_num - mean(age_num))

@@

age_me_effect <- as_tibble(ranef$familiar, rownames = "familiar")%>%
  mutate(grand_intercept = fixef%>%filter(term=="Intercept")%>%pull(Estimate),
         grand_slope = fixef%>%filter(term=="z_age")%>%pull(Estimate))%>%
  group_by(familiar) %>%
  expand(Estimate.Intercept,Estimate.z_age,grand_intercept,grand_slope,z_age = me_data$age_num - mean(me_data$age_num), mean_age = mean(me_data$age_num))%>%
  mutate(me_effect = plogis(grand_intercept + Estimate.Intercept+(Estimate.z_age+grand_slope)*z_age), 
         age = z_age+mean_age,
         item = familiar)%>%
  select(item, me_effect, age)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

@@

```{r predictions pragmatic model ex 3 adults}
comb_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, combPred , sep='\n'),
  data =list(data = comb_model_data),
  data_var = "dataFromR"
)
# 
# #saveRDS(inv_model_pred, "../saves/inv_model_pred.rds")
# #inv_model_pred <- readRDS("../saves/inv_model_pred.rds")
# 
comb_model_predictions <- comb_model_pred %>%
  separate(`0`, into = c("alignment","item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num))

@@

```{r}
ggplot(comb_model_predictions, aes(x=age,y = me_effect, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Combination Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  facet_grid(~alignment)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()
```

>>>>

var lexicon1 = function(utterance, obj, sem_knowledge){
  utterance.label == "novel_word" ? obj.shape == "novel_object" :
  utterance.label == "familiar_word" ? flip(sem_knowledge) ? 
    obj.shape == "familiar_object" :  
  flip() ? obj.shape == "familiar_object" : obj.shape == "novel_object" : 
  true
}

var lexicon2 = function(utterance, obj, sem_knowledge){
  utterance.label == "novel_word" ? obj.shape == "familiar_object" :
  utterance.label == "familiar_word" ? flip(sem_knowledge) ? 
    obj.shape == "familiar_object" : 
  flip() ? obj.shape == "familiar_object" : obj.shape == "novel_object" : 
  true
}

@@

```{r rsa model}
rsaModel <- '
var literalListener = cache(function(utterance, priorProbs, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior); 
    var lexicon = lexiconObject[lexiconName];
    var obj = sample( Categorical({vs: all_objects, ps: [.5,.5]}));
    if ("label" in utterance) {
      var truthValue = lexicon(utterance, obj, sem_knowledge);
      condition(truthValue)
    }
    return obj.shape 
  }})
}, 10000)

@@

  var speakerOptimalityParameters = {
    intercept: uniformDrift({a: -3, b: 3, width: 0.5}),
    slope: uniformDrift({a: 0, b: 3, width: 0.5})
  }

@@

# sem_know_model<- webppl(
#   program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
#   data = list(data = me_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 10,
#   cores = 10,
#   inference_opts = list(method = "incrementalMH",
#                         samples = 4000,
#                         burn = 2000,
#                         verbose = T,
#                         lag = 1,
#                         verboseLag = 500)
# )

# load this for model parameters
sem_know_model <- readRDS("../saves/sem_know_new_model.rds")#%>%

# load this for model parameters with wider prior on so_slope
#saveRDS(sem_know_model, "../saves/sem_know_new_model_2.rds")

@@

```{r}
#ggsave("../graphs/item_params_by_chain.pdf", width = 8, height = 3, scale = 1.5)
```

@@

```{r}
# me_model_data <- read_csv("../data/me.csv")%>%
#     filter(trial != "train1",
#            trial != "train2")%>%
#   mutate(age = age_num - min(age_num))
# 
# sem_know_run_summary <- item_params%>%
#   spread(parameter,value)%>%
#   group_by(Chain,item)%>%
#   summarise(overall_int = estimate_mode(int),
#             overall_slope = estimate_mode(slope))
# 
# 
# sem_know_run_iteration <- item_params%>%
#   spread(parameter,value)
# 
# sem_know_chain <- sem_know_run_iteration %>%
#   left_join(sem_know_run_summary)
# 
# chain_comp <- sem_know_chain %>%
#   left_join(expand.grid(
#       age = unique(me_model_data$age),
#       item = unique(me_model_data$familiar)), .) %>%
#   mutate(sem_know_ci = plogis(int + slope * age),
#          sem_know = plogis(overall_int + overall_slope * age)) %>%
#   group_by(Chain,item, age) %>%
#   summarise(sem_know = estimate_mode(sem_know), 
#             sem_know_uci = hdi_upper(sem_know_ci),
#             sem_know_lci = hdi_lower(sem_know_ci))
# 
# ggplot(chain_comp, aes(x = age+2, y= sem_know, fill = factor(Chain)))+
#   geom_line(size = .3, aes(col = factor(Chain)))+
#   geom_ribbon(aes(ymin=sem_know_lci, ymax = sem_know_uci), alpha = .05)+
#   ylab("Semantic knowledge")+
#   xlab("Age")+
#   ylim(0,1)+
#   facet_wrap(~item, nrow = 3)+
#   theme_few()+
#   scale_colour_discrete()+
#   scale_fill_discrete()

#ggsave("../graphs/model_by_chain.pdf", width = 5, height = 4, scale = 1.5)
```

@@

```{r}
# me_model_data <- read_csv("../data/me.csv")%>%
#     filter(trial != "train1",
#            trial != "train2")%>%
#   mutate(age = age_num - min(age_num))
# 
# so_run_iteration <- global_params%>%
#   spread(parameter, value)
# 
# so_run_summary <- so_run_iteration%>%
#   group_by(Chain)%>%
#   summarise(overall_int = estimate_mode(int),
#             overall_slope = estimate_mode(slope))
# 
# so_run <- so_run_iteration %>%
#   left_join(so_run_summary)
# 
# so_run_comp <-so_run%>%
#   group_by(Chain,Iteration)%>%
#   tidyr::expand(int,slope,overall_int,overall_slope,age = unique(me_model_data$age))%>%
#   mutate(so_ci = int + slope * age,
#          so = overall_int + overall_slope * age) %>%
#   group_by(Chain, age) %>%
#   summarise(so = estimate_mode(so), 
#             so_uci = hdi_upper(so_ci),
#             so_lci = hdi_lower(so_ci))
# 
# ggplot(so_run_comp, aes(x = age+2, y= so, fill = factor(Chain)))+
#   geom_line(size = .4, aes(col = factor(Chain)))+
#   geom_ribbon(aes(ymin=so_lci, ymax = so_uci), alpha = .05)+
#   ylab("speaker optimality")+
#   xlab("Age")+
#   #facet_grid(~run)+
#   theme_few()+
#   scale_colour_discrete()+
#   scale_fill_discrete()
# 
# ggsave("../graphs/so_by_chain.pdf", width = 4, height = 3, scale = 1.5)
```

@@

# sem_know_cor_params  <- readRDS("../saves/run_correlations.rds")%>%
#   separate(Parameter, into = c("parameter", "type","item"), sep = ",")%>%
#   filter(parameter == "parameter", type == "items")%>%
#   separate(value, into = c("int", "slope"),sep = ", ")%>%
#   mutate(int = as.numeric(str_remove_all(int, "[(c]")),
#          slope = as.numeric(str_remove(slope, "[)]")))%>%
#   gather(parameter, value, int, slope)
# 
# 
# sem_know_cor_params <- sem_know_cor_params%>%
#   filter(Chain != "15",
#          Chain != "1",
#          Chain != "30",
#          Chain != "6",
#          Chain != "5",
#          Chain != "7",
#          Chain != "3")

@@

```{r}
ggplot(item_params%>%filter(parameter == "int"), aes(x = value))+
  geom_density(alpha = 0.3)+
  xlab("Parameter")+
  facet_grid(Chain~item, scales = 'free')+
  theme_few()

@@

```{r}
item_params_cor <- item_params %>%
  group_by(Chain,item, parameter)%>%
  summarise(mode = estimate_mode(value))
```

@@

## For comparison: GLMM for ME effect and data
```{r}
me_data <- read_csv("../data/me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(z_age = age_num - mean(age_num))

@@

age_me_effect <- as_data_frame(ranef$familiar, rownames = "familiar")%>%
  mutate(grand_intercept = fixef%>%filter(term=="Intercept")%>%pull(Estimate),
         grand_slope = fixef%>%filter(term=="z_age")%>%pull(Estimate))%>%
  group_by(familiar) %>%
  tidyr::expand(Estimate.Intercept,Estimate.z_age,grand_intercept,grand_slope,z_age = me_data$age_num - mean(me_data$age_num), mean_age = mean(me_data$age_num))%>%
  mutate(me_effect = plogis(grand_intercept + Estimate.Intercept+(Estimate.z_age+grand_slope)*z_age), 
         age = z_age+mean_age,
         item = familiar)%>%
  select(item, me_effect, age)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

@@

```{r predictions pragmatic model ex 3 adults}
comb_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, combPred , sep='\n'),
  data =list(data = comb_model_data),
  data_var = "dataFromR"
)
# 
# #saveRDS(inv_model_pred, "../saves/inv_model_pred.rds")
# #inv_model_pred <- readRDS("../saves/inv_model_pred.rds")
# 
comb_model_predictions <- comb_model_pred %>%
  separate(`0`, into = c("alignment","item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num),
         source = "Model")

@@

```{r}
p_comb_model <- ggplot(comb_model_predictions, aes(x=age,y = me_effect, col = item))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Combination Model Predictions")+
  labs(x="Age",y="Proportion correct")+
  facet_grid(~alignment)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()
```

data_comb <- read_csv("../data/combination.csv") %>%
  filter(condition != "train")%>%
  mutate(alignment = condition,
         item = familiarObject)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(familiarObject), mean_aoa), 
         source = "Data")

p_comb_data <- ggplot(data = data_comb, aes(x = age_num, y = correct, col = item, fill = item)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  #geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F, alpha = .5)+
  geom_smooth(method = "loess", se = F, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(~alignment)+
  ggtitle("Combination Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d()+
  scale_fill_viridis_d()
```

```{r}
ggarrange(p_comb_model,p_comb_data, common.legend = T, legend = "right", nrow = 2, ncol = 1)
```

```{r}
#ggsave("../graphs/combination_model_data.pdf", width = 7, height = 7, scale = 1)
```

```{r}
ggplot(data = data_comb, aes(x = age_num, y = correct, col = source, fill = source)) +
  geom_hline(yintercept = 0.5, lty=2)+
  #geom_jitter(width = 0, height = 0.05, alpha = .6)+
  #geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F, alpha = .5)+
  geom_smooth(method = "loess", se = T, alpha = .5, span = 2)+
  geom_line(data = comb_model_predictions, aes(x=age,y = me_effect, col = source), size = 1)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  #facet_grid(condition~familiarObject)+
  facet_grid(alignment~item)+
  #ggtitle("Combination Data")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)
```

```{r}
#ggsave("../graphs/combination_model_data_by_item.pdf", width = 8, height = 2, scale = 1.5)
```

```{r}
binnned_data <- data_comb%>%
  group_by(subage, alignment, item)%>%
  summarize(k = sum(correct), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         data_mean = (a-1)/(a+b-2),
         data_ci_lower  = qbeta(.025, a, b),
         data_ci_upper = qbeta(.975, a, b),
         subage = factor(subage))%>%
  select(-a,-b,-n,-k)

binned_model <- comb_model_predictions %>%
  mutate(subage = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(subage, alignment, item)%>%
  summarise(model_mean = estimate_mode(me_effect),
            model_ci_lower = hdi_lower(me_effect),
            model_ci_upper = hdi_upper(me_effect))

cor_plot <- left_join(
  binnned_data,
  binned_model)

ggplot(data = cor_plot,aes(x = model_mean, y = data_mean, col = item)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .7)+
  geom_errorbarh(aes(xmin = model_ci_lower, xmax = model_ci_upper), height = 0,size = .7)+
  geom_point(size = 2,  stroke = 1)+
  coord_fixed()+
 stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = model_mean, y = data_mean), inherit.aes = F, size = 3)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  #ggtitle("3-year-olds")+
  facet_grid(~subage)+
  theme_few() + 
  scale_colour_viridis_d(name ="Item")
```

```{r}
#ggsave("../graphs/combination_correlation.pdf", width = 8, height = 4, scale = 1.2)
```
